### 2024-12-25 paper/header.tex
Add files via upload

% **************GENERATED FILE, DO NOT EDIT**************

\title{DMRGenie: Fully graphical user interface for entanglement renormalization}

\author[1]{Kiana Gallagher}
\author[1]{Aaron Dayton}
\author[2]{Drew Leske}
\author[2]{Sarah Huber}
\author[1,3,4]{Thomas E.~Baker}
\affil[1]{Department of Physics \& Astronomy, University of Victoria, Victoria, British Columbia V8P 5C2, Canada}
\affil[2]{Research Computing Services, University of Victoria, Victoria, British Columbia V8P 5C2, Canada}
\affil[3]{Department of Chemistry, University of Victoria, Victoria, British Columbia V8P 5C2, Canada}
\affil[4]{Centre for Advanced Materials and Related Technologies, University of Victoria, Victoria, British Columbia V8P 5C2, Canada}

\keywords{Julia, Entanglement Renormalization, Tensor Networks}

\hypersetup{
pdftitle = {My JuliaCon proceeding},
pdfsubject = {JuliaCon 2019 Proceedings},
pdfauthor = {1st author, 2nd author, 3rd author},
pdfkeywords = {Julia, Optimization, Game theory, Compiler},
}