### 2024-12-29 paper/header.tex
Add files via upload

% **************GENERATED FILE, DO NOT EDIT**************

\title{DMRGenie: Graphical interface for entanglement renormalization}

\author[1]{Kiana Gallagher}
\author[1]{Aaron Dayton}
\author[2]{Drew Leske}
\author[2]{Sarah Huber}
\author[134]{Thomas E. Baker}
\affil[1]{Department of Physics \& Astronomy, University of Victoria, Victoria, British Columbia V8P 5C2, Canada}
\affil[2]{Research Computing Services, University of Victoria, Victoria, British Columbia V8P 5C2, Canada}
\affil[3]{Department of Chemistry, University of Victoria, Victoria, British Columbia V8P 5C2, Canada}
\affil[4]{Centre for Advanced Materials and Related Technologies, University of Victoria, Victoria, British Columbia V8P 5C2, Canada}

\keywords{Julia, Genie, DMRjulia, TENPACK, Entanglement Renormalization, Tensor Networks, Graphical User Interface}

\hypersetup{
pdftitle = {DMRGenie: Graphical interface for entanglement renormalization},
pdfsubject = {JuliaCon 2022 Proceedings},
pdfauthor = {Kiana Gallagher, Aaron Dayton, Drew Leske, Sarah Huber, Thomas E. Baker},
pdfkeywords = {Julia, Genie, DMRjulia, TENPACK, Entanglement Renormalization, Tensor Networks, Graphical User Interface},
}